Group 2 Final Project Report (revB1218)

### 2013-12-19 reportLatex/paper.tex
flipped pref sections

\title{Group 2 Final Project Report (revA1218)}
\author{Jay Guild, Nathan Miller, Matthew L Weber}

@@

\section{Introduction}
\IEEEPARstart{A}{cceleration} of a cryptographic algorithms on an embedded device remains an outstanding challenge. New System on Chip (SoC) FPGA technologies provide more options to ofﬂoad or accelerate the cryptographic computations, however, integrating these technologies into existing designs can be challenging. Similarly, migrating designs between different SoC FPGA platforms can also require a signiﬁcant investment of time and resources. Even though the system functionality and underlying algorithms remain unchanged, the design teams are forced to re-invent their designs for each new platform.

This project sought to characterize the effort of porting a simple cryptographic algorithm to a SoC FPGA prototype containing a fabric based accelerator and provide empirical performance data showing the beneﬁts of doing so. Towards this end, we implemented the SHA-1 algorithm on a Xilinx Zynq SoC FPGA 702 development kit and compared the performance of our SHA-1 fabric to a pure software implementation running on the Zynq’s ARM core.  The Linux cryptodev framework provided transparent access to the SHA-1 function, permitting the user-space to be ignorant of where the SHA-1 function was actually being performed.   Our project uses similar concepts implemented on Texas Instruments SoC processors to tie their hardware accelerators into the Linux Kernel.

@@

\section{Why is this interesting...}
In the last two years various industry leading programmable logic companies have invested millions in creating consolidated system on a chip (SoC) products. Fusing mainstream processor technology (primarily ARM-based cores) with the latest fabric technology, the result has been the creation of highly conﬁgurable embedded devices. This is very relevant to bleeding edge signal processing developments (cryptography/video/audio/RF), where a general purpose processor is commonly paired with a programmable device for initial prototyping and in some cases the ﬁnal product. The latest SoC FPGA technology allows a consolidation that has largely eliminated many interface design activities and has encouraged more of a accelerator/coprocessor style of development. The key to this ﬂexibility is having the SoC execute code that may tap into a fabric logic to parallelize an operation or having SoC completely hand off data to the fabric (leaving the SoC free to perform other operations).

@@

\section{Design Description}
The design consists of the components outlined in Figure \ref{fig_archBlkDetailed}.  Each of these is expanded in more detail in the sections that follow below.
\subsection{Hardware}
The Xilinx Zynq SoC FPGA 702 development kit was used as our target hardware. We acquired one development kit for our project and coordinated integration efforts to share the workload of only having one kit.  ISE tool support was included in the university's 14.6 license and the ISE EDK builder was used to pull together the embedded design. The Zynq ARM core Linux maturity was adequate in the Xilinx 14.6 release and Xilinx had some pretty good examples of interfacing between the fabric and processor. \cite{XIL2013} \cite{XILGNU2013}   The Zynq consists of two parts, the processing system 7 (PS7) and the programmable logic (PL). The PL might also be referred to as the “fabric” in the sections that follow.  The Zynq development kit was interfaced to via Ethernet and serial. With a base ARM load-set executing a Linux Kernel conﬁgured for modular development and capable of loading the PL images.
\subsection{Software}
For the software operating environment a full embedded Linux 3.12 Kernel and filesystem were used, allowing use of existing userspace applications and the Linux kernel’s existing infrastructure for cryptography.\cite{WIKPE2013} \cite{OCF2013} This configuration allowed the project to maximize reuse of the baseline algorithms already present in open source software and focus on gathering benchmark comparison data of our hardware accelerated approach.  The Xilinx Zynq board ARM OS load-set was based on Xilinx GIT tree source [14], with the Kernel/Rootfs being compiled into loadable binaries using the Buildroot \cite{BUILDROOT2013} embedded build system. We also added the Linux Cryptodev 1.6 support to the kernel and utilized the Cryptodev sha speed test app as part of our Buildroot rootfs.  The first stage bootloader (FSBL) was created using the Xilinx 14.6 EDK toolflow to export a XSDK FSBL project for our custom PL design.
\subsubsection{TestApp}
The test application used for executing the benchmarking of the baseline software algorithms on the SoC was based on the sha1\_speed application provided by the cryptodev-linux package.  It performed a number of block size test scenarios and captured throughput performance of each.

@@

Finally, our implementation of the SHA1 algorithm on the Zynq PL relied upon the CPU to supply it with data.  It had no mechanism to read memory directly.  This additional overhead of going through the CPU gives a distinct advantage to the pure software approach, which also has to read the hash message from memory, but then doesn’t have to redirect the data onto the slow(er) AXI bus in order to calculate the hash result.
\subsection{Future Enhancements}
There are a number of enhancements we could make to our PL design to potentially improve its performance.

@@

One final improvement would be to drop the SHA-1 hash altogether, and pick a hashing algorithm that is more amenable to parallelism or involves a complicated series of computations on usual data sizes.  Additional research into hash algorithms would be needed if this approach was taken.
\subsection{Runtime analysis of the AXI bus transactions}
The chipscope screenshots necessary for the following discussions have been captured in Appendix B. The first scenario captures the initial software status checking and loading of the data to be hashed  (Figure B-3).  The read valid signal was used to capture the initial software access to the status register (Figure B-2).  Initially, the (Figure B-3) shows the softwave time to determine to begin writing data took 50uSec from the time it initially read the status register noting that the algorithm was ready for data until the data actually hits the bus.  Next the complete software load of the data took 970uSec.  Then the software wrote a value of 0x1 to the control register to tell the user interface to begin processing the data.  At this point the capture ends (only had a buffer to view ~1.4uSec of transactions).  The time between when data is loaded and then the hash is computed is purely based on the algorithm compute time at 50Mhz characterized in simulation.

The chipscope captures have been archived in our project repository under the chipscope folder.  We also included the cdc file and a chipscope setup file if it’s desired to actively review transactions.  The *.vcd files contain both an initial and final capture showing the software interfacing and algorithm entity inputs/outputs.  If a viewer is needed for the vcd files, we suggest http://www.iss-us.com/wavevcd/Wave\_msi.zip.  The setup file for this application is also checked in with our vcd files.  If the setup file doesn’t happen to work, open the respective vcd file and select the read/write address, data, and valids.  (They’re the signals with short names/labels.)

\section{Reproducing Results}
\subsection{Project Materials}
All of our project files submitted with this report are also stored in the following GIT repository (https://github.com/matthew-l-weber/cpre583\_pjt/).  The folder structure is separated into a folder for buildroot (ie the rootfs and kernel build system), a folder for our initial sha testbenches, a folder capturing our Chipscope files, and a folder for our toplevel XPS project that creates our Zynq image with the embedded pcore SHA1 module. 
\subsection{Simulation}
Software simulation was done on a Gnome Desktop, Red Hat, Linux system running Xilinx version 14.6 ISE and ModelSIM.  We used an emulator in bash mode and browse to the xpsTopLevelProj\_FinalProj/pcores/glue\_v1\_00\_a/devl/projnav folder in the files provided.  From here source the setup.sh file and then launch the ISE with the command ise glue.xise.  In ISE we verified all items were listed and have no conflicts.

@@

In our original project proposal we expressed our intent to use a high-level synthesis tool to a convert a C implementation of the SHA-1 algorithm directly into VHDL.  We were then going to compare our hand-coded core against the synthesized version to determine how it compared in terms of performance and resource utilization.  In our proposal we listed two tools:  Xilinx HLS and Jacquard ROCCC 2.0.  in the course of working on this project we discovered that the university did not have a license for Xilinx HLS.  We were not inclined to purchase a license ourselves for a single project (about \$2000).  This left the ROCCC 2.0 compiler.  Although the ROCCC web-site is operational, the project itself appears to be either defunct or simply uninterested in responding to students.  All inquires to get a student license for the ROCCC 2.0 compiler went unanswered.
\section{Conclusion}
The implementation of a PL SHA1 cryptographic algorithm using the Linux kernel Crypto API showed to be successful.  With modifications to the user logic PL interface that would make the system custom to our application, integration of a SHA1 PL core, and tying it all together in simulation and on hardware.  In this process we thought that the core initially could operate with a lower process time if the state machine was driven by interrupts instead of process control to allow the software to execute parallel processes instead of waiting for data.  We did not explore this avenue as it was not in the scope of the project.  To test the system we stepped through a software simulated validation of the pieces with a final test on sending multiple block messages that gave successful data.  Upon examining the waveforms in the simulation we discovered the process time in the SHA-1 core could be optimized a bit more by using a wider AXI bus to send more than one 32 bit word at a time for more parallelized processing.
With a successful implementation of a software simulated SHA-1 we moved to a hardware implementation on a Zynq board.  We used the same approach as the software validation and stepped through tests to verify that all pieces were working properly with the final test being a multi block message being passed through the core and validated on the output.  This proved to be successful in the end, but has significant performance depreciation in speed.  We attributed this to the 50 MHz clock signal on the Zynq board and AXI bus to the ARM core while the software version ran a 1 GHz clock signal and all three pieces were in phase on the clock signal.  We achieved all middle level milestones for the project while discovering several areas for future development and extended data for this report.

>>>>

\title{Group 2 Final Project Report (revB1218)}
\author{Jay Guild, Nathan Miller, Matthew L Weber}

@@

\section{Introduction}
\IEEEPARstart{A}{cceleration} of a cryptographic algorithms on an embedded device remains an outstanding challenge. New System on Chip (SoC) FPGA technologies provide more options to offload or accelerate the cryptographic computations, however, integrating these technologies into existing designs can be challenging. Similarly, migrating designs between different SoC FPGA platforms can also require a significant investment of time and resources. Even though the system functionality and underlying algorithms remain unchanged, the design teams are forced to re-invent their designs for each new platform.

This project sought to characterize the effort of porting a simple cryptographic algorithm to a SoC FPGA prototype containing a fabric based accelerator and provide empirical performance data showing the benefits of doing so. Towards this end, we implemented the SHA-1 algorithm on a Xilinx Zynq SoC FPGA 702 development kit and compared the performance of our SHA-1 fabric to a pure software implementation running on the Zynq’s ARM core.  The Linux cryptodev framework provided transparent access to the SHA-1 function, permitting the user-space to be ignorant of where the SHA-1 function was actually being performed.   Our project uses similar concepts implemented on Texas Instruments SoC processors to tie their hardware accelerators into the Linux Kernel.

@@

\section{Why is this interesting...}
In the last two years various industry leading programmable logic companies have invested millions in creating consolidated system on a chip (SoC) products. Fusing mainstream processor technology (primarily ARM-based cores) with the latest fabric technology, the result has been the creation of highly configurable embedded devices. This is very relevant to bleeding edge signal processing developments (cryptography/video/audio/RF), where a general purpose processor is commonly paired with a programmable device for initial prototyping and in some cases the final product. The latest SoC FPGA technology allows a consolidation that has largely eliminated many interface design activities and has encouraged more of a accelerator/coprocessor style of development. The key to this flexibility is having the SoC execute code that may tap into a fabric logic to parallelize an operation or having SoC completely hand off data to the fabric (leaving the SoC free to perform other operations).

@@

\section{Design Description}
The design consists of the components outlined in Figure \ref{fig_archBlkDetailed}.  Each of these is expanded in more detail in the sections that follow below.
\subsection{Hardware}
The Xilinx Zynq SoC FPGA 702 development kit was used as our target hardware. We acquired one development kit for our project and coordinated integration efforts to share the workload of only having one kit.  ISE tool support was included in the university's 14.6 license and the ISE EDK builder was used to pull together the embedded design. The Zynq ARM core Linux maturity was adequate in the Xilinx 14.6 release and Xilinx had some pretty good examples of interfacing between the fabric and processor. \cite{XIL2013} \cite{XILGNU2013}   The Zynq consists of two parts, the processing system 7 (PS7) and the programmable logic (PL). The PL might also be referred to as the “fabric” in the sections that follow.  The Zynq development kit was interfaced to via Ethernet and serial. With a base ARM load-set executing a Linux Kernel configured for modular development and capable of loading the PL images.
\subsection{Software}
For the software operating environment a full embedded Linux 3.12 Kernel and filesystem were used, allowing use of existing userspace applications and the Linux kernel’s existing infrastructure for cryptography.\cite{WIKPE2013} \cite{OCF2013} This configuration allowed the project to maximize reuse of the baseline algorithms already present in open source software and focus on gathering benchmark comparison data of our hardware accelerated approach.  The Xilinx Zynq board ARM OS load-set was based on Xilinx GIT tree source [14], with the Kernel/Rootfs being compiled into loadable binaries using the Buildroot \cite{BUILDROOT2013} embedded build system. We also added the Linux Cryptodev 1.6 support to the kernel and utilized the Cryptodev sha speed test app as part of our Buildroot rootfs.  The first stage bootloader (FSBL) was created using the Xilinx 14.6 EDK toolflow to export a XSDK FSBL project for our custom PL design.
\subsubsection{TestApp}
The test application used for executing the benchmarking of the baseline software algorithms on the SoC was based on the sha1\_speed application provided by the cryptodev-linux package.  It performed a number of block size test scenarios and captured throughput performance of each.

@@

Finally, our implementation of the SHA1 algorithm on the Zynq PL relied upon the CPU to supply it with data.  It had no mechanism to read memory directly.  This additional overhead of going through the CPU gives a distinct advantage to the pure software approach, which also has to read the hash message from memory, but then doesn't have to redirect the data onto the slow(er) AXI bus in order to calculate the hash result.
\subsection{Runtime analysis of the AXI bus transactions}
The chipscope screenshots necessary for the following discussions have been captured in Appendix B. The first scenario captures the initial software status checking and loading of the data to be hashed  (Figure B-3).  The read valid signal was used to capture the initial software access to the status register (Figure B-2).  Initially, the (Figure B-3) shows the software time to determine to begin writing data took 50uSec from the time it initially read the status register noting that the algorithm was ready for data until the data actually hits the bus.  Next the complete software load of the data took 970uSec.  Then the software wrote a value of 0x1 to the control register to tell the user interface to begin processing the data.  At this point the capture ends (only had a buffer to view ~1.4uSec of transactions).  The time between when data is loaded and then the hash is computed is purely based on the algorithm compute time at 50Mhz characterized in simulation.

The chipscope captures have been archived in our project repository under the chipscope folder.  We also included the cdc file and a chipscope setup file if it’s desired to actively review transactions.  The *.vcd files contain both an initial and final capture showing the software interfacing and algorithm entity inputs/outputs.  If a viewer is needed for the vcd files, we suggest http://www.iss-us.com/wavevcd/Wave\_msi.zip.  The setup file for this application is also checked in with our vcd files.  If the setup file doesn't happen to work, open the respective vcd file and select the read/write address, data, and valids.  (They’re the signals with short names/labels.)
\subsection{Future Enhancements}
There are a number of enhancements we could make to our PL design to potentially improve its performance.

@@

One final improvement would be to drop the SHA-1 hash altogether, and pick a hashing algorithm that is more amenable to parallelism or involves a complicated series of computations on usual data sizes.  Additional research into hash algorithms would be needed if this approach was taken.
\section{Reproducing Results}
\subsection{Project Materials}
All of our project files submitted with this report are also stored in the following GIT repository (https://github.com/matthew-l-weber/cpre583\_pjt/).  The folder structure is separated into a folder for buildroot (ie the rootfs and kernel build system), a folder for our initial sha testbenches, a folder capturing our Chipscope files, and a folder for our toplevel XPS project that creates our Zynq image with the embedded pcore SHA1 module. 
\subsection{Simulation}
Software simulation was done on a Gnome Desktop, Red Hat, Linux system running Xilinx version 14.6 ISE and ModelSIM.  We used an emulator in bash mode and browse to the xpsTopLevelProj\_FinalProj/pcores/glue\_v1\_00\_a/devl/projnav folder in the files provided.  From here source the setup.sh file and then launch the ISE with the command ise glue.xise.  In ISE we verified all items were listed and have no conflicts.

@@

In our original project proposal we expressed our intent to use a high-level synthesis tool to a convert a C implementation of the SHA-1 algorithm directly into VHDL.  We were then going to compare our hand-coded core against the synthesized version to determine how it compared in terms of performance and resource utilization.  In our proposal we listed two tools:  Xilinx HLS and Jacquard ROCCC 2.0.  in the course of working on this project we discovered that the university did not have a license for Xilinx HLS.  We were not inclined to purchase a license ourselves for a single project (about \$2000).  This left the ROCCC 2.0 compiler.  Although the ROCCC web-site is operational, the project itself appears to be either defunct or simply uninterested in responding to students.  All inquires to get a student license for the ROCCC 2.0 compiler went unanswered.
\section{Conclusion}
The implementation of a PL SHA1 cryptographic algorithm using the Linux kernel Crypto API showed to be successful.  With modifications to the user logic PL interface that would make the system custom to our application, integration of a SHA1 PL core, and tying it all together in simulation and on hardware.  In this process we thought that the core initially could operate with a lower process time if the state machine was driven by interrupts instead of process control to allow the software to execute parallel processes instead of waiting for data.  We did not explore this avenue as it was not in the scope of the project.  To test the system we stepped through a software simulated validation of the pieces with a final test on sending multiple block messages that gave successful data.  Upon examining the waveforms in the simulation we discovered the process time in the SHA1 core could be optimized a bit more by using a wider AXI bus to send more than one 32 bit word at a time for more parallelized processing.
With a successful implementation of a software simulated SHA1 we moved to a hardware implementation on a Zynq board.  We used the same approach as the software validation and stepped through tests to verify that all pieces were working properly with the final test being a multi block message being passed through the core and validated on the output.  This proved to be successful in the end, but has significant performance depreciation in speed.  We attributed this to the 50 MHz clock signal on the Zynq board and AXI bus to the ARM core while the software version ran a 1 GHz clock signal and all three pieces were in phase on the clock signal.  We achieved all middle level milestones for the project while discovering several areas for future development and extended data for this report.
